\section{Introduction}\label{sec:intro}

Reference-conditioned identity preservation -- feed a reference portrait,
get the same character in a new scene -- is normally bought one of two
ways: train a dedicated adapter on a large paired dataset
\citep{ye2023ipadapter,li2023photomaker,wang2024instantid}, or switch to
an instruction-following edit model whose base was already pretrained
for image-to-image tasks \citep{blackforest2025fluxkontext}. Neither
route is available when the base model must remain a frozen, pure
text-to-image (T2I) diffusion transformer: no per-subject or general
fine-tuning, no departure from T2I semantics into edit-model territory,
no library modification. We show such a model -- a frozen anime T2I DiT,
Z-Image \citep{zimage2025} -- can be activated as a reference-conditioned
identity generator by two cheap, mechanistically independent
interventions: concatenating the reference image's attention key/value
tensors into a restricted band of self-attention layers at inference
time (zero training), and a rank-8 LoRA on 0.06\% of backbone parameters
within the same layer band (minutes of training, not per-subject).

Both mechanisms reach that parity, which lets us ask a sharper question
than ``does it work'': \emph{where} in the frozen transformer does the
one-shot identity signal live, and can it be localized precisely enough
to repair the one defect that dogs it? Within every handle we can test,
it cannot -- the signal is not localized. Identity-match rate is
statistically flat across four transformer-layer bands (a null that
replicates on a second held-out subject set and holds for \emph{both} the
injection and the trained mechanism), across attention-head subsets
chosen by reference affinity (matched-count controls), and across
denoising-timestep gates. The persistent tiled/``collage'' defect is
entangled with this distributed identity rather than separable from it:
five targeted repairs -- background isolation, head-gating, high-norm
reference-token suppression, late-timestep gating, and an
identity-supervised training loss -- each with a matched control, every
one trades identity away for coherence instead of removing the defect.
This is what a distributed, composition-entangled representation
predicts, and it is the paper's spine.

We learned to measure this honestly only after two reversals, themselves
a contribution. A frequency-gated fix for the injection mechanism's
copy-paste ``ghosting'' -- caused by the reference and target token grids
sharing one RoPE coordinate origin -- raises raw identity-match on a
development set (23\%$\to$40\%) but, under a defect-aware held-out judge,
also raises composition-collapse from 20\% to 82\%, dropping usable
output from 53\% to 15\%; identity-match alone would have shipped the
worse recipe. And a pre-registered band ablation's apparent 8.6$\times$
defect-rate gradient, CI-decisive on one set, failed our own audit -- no
cross-set replication, and a confound-free re-run of its reference arm
erases it. Identity-match-only, single-set evaluation is systematically
optimistic for this class of method.

\paragraph{Contributions.}
\textbf{(1) Cheap activation with a reference-driven control.} A frozen
pure-T2I DiT becomes a one-shot identity conditioner two mechanistically
independent, low-cost ways -- inference-time attention key/value
injection (zero training) and a 0.06\%-parameter band-LoRA -- each
reaching identity parity with a same-class trained adapter; a
shuffled-reference control confirms the conditioning is genuinely
reference-driven rather than a prompt-only floor, and exposes that a
stronger baseline's identity number is itself partly such a floor.
\textbf{(2) An anatomy of the identity signal (the core claim).}
One-shot identity in this frozen DiT behaves as a \emph{distributed,
composition-entangled} attention code: identity-match is flat across
layer bands (cross-set-replicated, and concurring across the injection
and the trained mechanism), across reference-affinity head subsets, and
across denoising-timestep gates -- a locality we do not find, in contrast
to layer-locality assumptions in prior injection work
\citep{zhang2025freecus} and head-specialization assumptions in
attention-routing methods (whose priors, we note, may simply not transfer
to this frozen-T2I anime setting rather than being refuted by it). We
give a RoPE-coordinate mechanism for the ghosting failure and show the
collage defect and identity co-form early in denoising; the resulting
prediction -- that no localized repair separates them -- is borne out by
five matched-control repairs that all trade identity for coherence.
\textbf{(3) A diagnostic and an evaluation protocol.} A
peaked-versus-diffuse attention-selectivity signature, observed across
two architectures and offered as a diagnostic rather than a law,
separates a model that transfers structured identity from one that
transfers only mean color; and a paired identity-and-defect, cross-set
held-out protocol -- motivated by two measurement reversals that
identity-only, single-set scoring would have shipped -- that catches
recipe- and protocol-level optimism such scoring misses.
