\section{Related Work}\label{sec:related}

\paragraph{Training-free attention/KV injection.} MasaCtrl
\citep{cao2023masactrl}, Cross-Image Attention \citep{alaluf2023crossimage},
StyleAligned \citep{hertz2023stylealigned}, ConsiStory
\citep{tewel2024consistory}, FreeCus \citep{zhang2025freecus}, Personalize
Anything \citep{feng2025personalizeanything}, and FreeGraftor
\citep{yao2025freegraftor} all report the failure mode we do -- reference
layout leaking into the target -- and patch it differently: layer
restriction, AdaIN, correspondence warping, zero-position concatenation,
coordinate rebinding. All but the last two target real-photo UNets/DiTs;
to our knowledge we are the first to apply this family to a \emph{pure
anime T2I DiT}.

\paragraph{Trained adapters.} IP-Adapter \citep{ye2023ipadapter} (our
baseline below), PhotoMaker \citep{li2023photomaker},
and InstantID \citep{wang2024instantid} condition on a pooled image
embedding; AnimeAdapter \citep{han2026animeadapter}, the only
anime-native zero-shot method we found, has no public weights. We also
run \texttt{krea2-identity-edit} (community LoRA, 12.9B general-domain
Krea-2-Turbo) as a second baseline: it injects the source latent as a
RoPE frame-1 in-context token rather than a pooled embedding.

\paragraph{RoPE position manipulation.} OminiControl
\citep{tan2024ominicontrol}, UNO \citep{wu2025uno}, EasyControl
\citep{zhang2025easycontrol}, and FLUX.1 Kontext
\citep{blackforest2025fluxkontext} all offset or strip RoPE from
conditioning tokens -- close to folk knowledge for DiTs by 2026.
Untwisting RoPE \citep{mikaeili2026untwisting} is the closest neighbor to
our frequency-gating idea but scales frequencies continuously rather than
hard-cutting; its own ablation (low-frequency-only scaling fails)
indirectly supports our finding that high-frequency suppression is the
operative mechanism. ReRoPE \citep{li2026rerope} nulls a frequency band on
a video DiT's temporal axis for a diagnostic ablation, not reference
conditioning -- a different setting.

\paragraph{Small-LoRA in-context conditioning.} ICEdit's diptych LoRA
($\sim$1\% of parameters, fill-pretrained base) and UNO/OminiControl show
a small LoRA can teach in-context conditioning to a DiT already
pretrained for image-to-image tasks. Our training-light mechanism asks
whether the same recipe transfers to a backbone with \emph{no} such
pretraining, at 0.06\% of parameters, and reports a pre-registered
ablation whose primary hypothesis fails rather than only the
configuration that worked.

\paragraph{Anime identity metrics.} CCIP (\texttt{imgutils}) is used
throughout prior anime-generation tooling but, per our search, never
cited or benchmarked in a paper; the ArcFace/FaceNet-style recognizers
IP-Adapter's and InstantID's own evaluations use do not detect faces on
anime images at all. We treat CCIP as a pre-filter only and report VLM
(Gemini 3.1 Pro) chain-of-thought judgments -- identity match
\emph{and} visible defects jointly -- as the final signal.
