\section{Experiments: Training-Free Mechanism}\label{sec:results-tf}

\subsection{Evaluation protocol}\label{sec:protocol}

CCIP (anime-domain character similarity, SAME threshold 0.178) is a fast
pre-filter only; it is confirmed insensitive to several failure modes
(hair-color changes, gender/hairstyle flips) a downstream VLM judge
catches. We report Gemini 3.1 Pro's \texttt{same\_character: yes/no/partial}
judgment (ref vs.\ generated, ignoring pose/outfit/background) as the
primary identity metric. Held-out and baseline judgments add a second,
independent \texttt{visible\_defect: yes/no} field to the same judge call
(rendering artifacts only -- tiling, duplication, broken anatomy;
explicitly not art-style or pose) and report the \emph{joint} metric,
using an explicit chain-of-thought prompt (the model reasons through each
identity axis and the composition check in prose before emitting JSON).

\subsection{Dev-set ablation (band width $\times$ position gating)}\label{sec:devablation}

10 subjects, 2--4 seeds, random sample, $n{=}20$--$40$, Wilson 95\% CI.
Every row is measured on the same 10 subjects band width and
\texttt{lowfreq\_k} were tuned against -- reported to document the tuning
process; the held-out validation below is the number that matters for
any external claim.

\begin{table}[t]
\centering\small
\begin{tabular}{@{}lrrr@{}}
\toprule
Recipe & $n$ & yes & no (95\% CI) \\
\midrule
\texttt{strip}, band 21--25 (prior default) & 40 & 18\% & 25\% [14,40] \\
\texttt{center}, band 21--25 & 20 & 10\% & 55\% [34,74] \\
\texttt{lowfreq} $k{=}8$, band 21--25 & 20 & 0\% & 30\% [14,52] \\
\texttt{lowfreq} $k{=}8$, band 10--25 & 40 & 58\% & \textbf{0\% [0,9]} \\
\bottomrule
\end{tabular}
\caption{Dev-set band/position ablation (identity-only metric).}
\label{tab:devablation}
\end{table}

\subsection{Held-out validation: the recipe reverses}\label{sec:heldout}

Recipe frozen at \texttt{band=10-25, refpos=lowfreq, lowfreq\_k=8,
sched=taper, w=4.0} \emph{before} running this section. 30 subjects
disjoint from the dev set, 2 seeds, paired against the old default on the
same subjects/seeds, single blind judgment pass.

\begin{table}[t]
\centering\small
\begin{tabular}{@{}lrrrr@{}}
\toprule
Recipe & $n$ & defect & clean-yes & clean-usable \\
\midrule
old (\texttt{strip}) & 60 & 20\% [12,32] & 13\% [7,24] & \textbf{53\% [41,65]} \\
new (\texttt{lowfreq}) & 60 & \textbf{82\% [70,89]} & 8\% [4,18] & 15\% [8,26] \\
\bottomrule
\end{tabular}
\caption{Held-out defect and joint (usable-output) metric. clean-yes = identity=yes \& no defect; clean-usable = identity$\in\{$yes,partial$\}$ \& no defect.}
\label{tab:heldout}
\end{table}

Read in isolation, the marginal identity metric looks like a modest win
for the new recipe (yes: 23\%$\to$40\%) -- exactly the framing the
defect field above was designed to guard against. 79\% of the new
recipe's ``yes'' verdicts also carry a defect flag
(19/24). On usable-output rate the 95\% CIs do not overlap: the
position-free baseline is usable 53\% of the time; the frequency-gated
recipe, 15\%. The defect is a composition failure -- collaged/tiled
sub-images, picture-in-picture insets, duplicated fragments
(Fig.~\ref{fig:holdout}) -- far more frequent than a minor artifact. Both
arms share the same subjects, reference images, and seeds; only the
injection recipe differs, ruling out a data-source confound. The
reversal is CI-decisive on defect rate and clean-usable rate; on strict
identity match alone the CIs overlap (directional, not decisive, at
$n{=}60$/arm). \textbf{Conclusion: \texttt{strip}/band 21--25 is the
recipe this work recommends for deployment}; \texttt{lowfreq}/band
10--25 remains a research direction with a real, unresolved defect mode.

\subsection{Baseline comparison}\label{sec:baseline}

IP-Adapter-SDXL (\texttt{ip-adapter-plus-face\_sdxl\_vit-h}, no per-subject
training, same 30 held-out subjects) is the same-\emph{method}-class
trained-adapter baseline -- its own SDXL backbone, not Z-Image, and never
adapted to the anime domain it is tested on here, so this is a same-class
reference point, not an architecture- or domain-controlled comparison.
\texttt{krea2-identity-edit} (community LoRA on the 12.9B general-domain
Krea-2-Turbo, zero retraining by us) is likewise purpose-trained and
architecturally distinct. Neither baseline is architecture-matched to our
Z-Image backbone; both are the strongest publicly available bars we could
run at zero additional training cost.

\begin{table}[t]
\centering\small
\begin{tabular}{@{}lrrr@{}}
\toprule
Recipe & defect & clean-yes & clean-usable \\
\midrule
IP-Adapter-SDXL & 33\% & 18\% [11,30] & 58\% [46,70] \\
krea2-identity-edit & \textbf{2\% [0,9]} & \textbf{33\% [23,46]} & 58\% [46,70] \\
this work, \texttt{strip} & 20\% [12,32] & 13\% [7,24] & 53\% [41,65] \\
this work, \texttt{lowfreq} & \textbf{82\% [70,89]} & 8\% [4,18] & 15\% [8,26] \\
\bottomrule
\end{tabular}
\caption{Four-arm held-out comparison, $n{=}60$ each.}
\label{tab:baseline}
\end{table}

Restricting to the two general-adapter-class arms (\texttt{strip},
IP-Adapter): both metrics nominally favor IP-Adapter, but every CI
overlaps heavily -- neither ranking is decisive at this sample size. The
defensible claim is \textbf{training-free \texttt{strip} is statistically
indistinguishable from a same-class trained adapter's strict-identity
rate, at zero training cost} -- not that it beats one. That does not
extend to krea2: against \texttt{strip}, krea2 is CI-decisive on defect
rate and clean-yes, but only directional on clean-usable. krea2's
near-elimination of the collage defect fits its mechanism -- a source
latent trained end-to-end as a structured in-context token, unlike raw
K/V concatenated post-hoc into attention pools the frozen model never
expected -- but krea2 also has the \emph{highest} outright-miss rate of
all four arms (40\% vs.\ IP-Adapter's 15\%): most reliable when it
works, most likely to miss outright when it does not. Manual inspection
(Fig.~\ref{fig:baseline}) shows three qualitatively distinct failure
signatures -- \texttt{lowfreq} collages, IP-Adapter garbles hands, krea2
normalizes/drops accessories toward a generic prior rather than
mis-rendering them -- not one mode at three severities. \texttt{lowfreq}
remains decisively worst on both strict metrics despite the highest raw
identity-match rate, reinforcing that
identity-match alone does not reliably rank these methods.

\subsection{No localized handle for the defect}\label{sec:repairs}

If the identity signal were localized, the collage defect should have a
local handle -- a layer, a head, a token, or a denoising phase -- that
suppresses it while sparing identity. We probed four such handles on the
injection mechanism, each against a matched control, and the training
objective of Sec.~\ref{sec:auxloss} is a fifth; none separates the defect
from identity. \emph{(i) Background isolation.} Matting the reference to
a white background (the content-leakage hypothesis) leaves the
composition defect essentially unchanged (combined collage/other rate
$88\%\to80\%$; the residual is a top-left positional artifact class, not
per-reference leakage). \emph{(ii) Head-gating.} Restricting injection to
the top-quarter attention heads by reference affinity does not beat a
random-quarter or bottom-quarter control ($n{=}12$), and the
top-affinity selection is in fact weakest on identity, so the apparent
``fewer heads, cleaner'' effect is injection \emph{count}, not head
selection. \emph{(iii) Outlier-token suppression.} Zeroing high-norm
reference tokens is indistinguishable from a same-count random-drop
control. \emph{(iv) Late-timestep gating.} Withholding injection until
the low-noise phase removes the collage cleanly, but identity collapses
with it (0/5 subjects retain reference-specific identity), because
identity and composition co-form in the early denoising pass. Each
result is individually modest, some at small $n$, but they agree in
direction across four independent injection-side handles plus the
training objective -- the outcome a distributed, composition-entangled
identity predicts and a localized one does not. We do not claim no handle
\emph{can} exist; we report that none of the natural layer-, head-,
token-, timestep-, or objective-level handles we tested works.
