\section{Limitations}\label{sec:limits}

\texttt{lowfreq} (band 10--25) is \textbf{not production-ready}: 82\%
held-out defect rate collapses usable-output to 15\% despite a higher
raw identity-match rate. \emph{Why} is partially decomposed: a
position-information-free variant at the same widened band is
statistically as defective (75\% vs.\ 71\% on a dev probe), ruling out
position-encoding precision as the driver -- injecting reference
content into layers 10--20 at all is what destabilizes; no cheap fix
emerged.
Geometric attributes (face/eye shape) remain a residual failure for the
recommended (\texttt{strip}) recipe even where composition is clean
(45\% partial rate). The frequency-gating sweet spot ($k{=}8$) and band
width (10--25) were tuned on a 10-subject dev set using a metric later
shown insufficient; re-tuning on a fresh, disjoint probe set is future
work, as are AnimeAdapter and FreeGraftor/UNO baselines (no public
weights / gated dependency).

For the training-light mechanism, $n{=}60$ is 30 subjects $\times$ 2
seeds, not 60 independent draws; headline comparisons therefore carry
subject-clustered CIs alongside per-image ones. The defect-gradient
audit chain (different-family cross-judge, seeded random flag audit,
cross-set replication, confound-free re-run) is reported in the results
section; each step independently weakened the gradient before the
ensemble refuted it, and the first warning was visible early -- GPT-5.5
had already failed to reproduce the gradient's magnitude. Attribution
of the confounded mid arm's advantage is now closed, not bounded: the
completed checkpoint$\times$prompt-prefix factorial shows the prefix
effect is checkpoint-specific, not portable ($+13.3$pp within
run\_mix1, $+0$pp within run\_mix2), and the only clustered-CI-disjoint
marginal (checkpoint at prefix-on, $+26.7$pp) still overlaps at
prefix-off -- inseparable there from ordinary run-to-run training
variance. Absolute numbers are modest
regardless of arm: 51.7--55.0\% clean-usable / ${\sim}20$\% clean-yes
is not state-of-the-art accuracy in either line of this work; the
contribution is the anatomy of the identity signal and the evaluation
methodology, not benchmark accuracy. All experiments ran on a single
141GB-class GPU (LoRA training ${\sim}4$--$6.5$h per arm; generation
${\sim}35$\,min per 60 images).

\paragraph{Scope of the distributed-identity claim.} The
non-localization evidence is bounded by the handles we tested (continuous
layer bands, reference-affinity head subsets, high-norm token sets,
timestep gates); a finer or compositional handle could in principle
localize identity, and we claim only that these natural handles do not.
The two mechanisms are independent as \emph{interventions} but share
evaluation \emph{instruments} (one judge, one CCIP stack, one anime data
family), so their agreement is corroborative, not fully independent. The
cross-architecture selectivity signature is correlational on $n{=}2$
models -- a within-model RoPE-$\theta$ test was inconclusive by
construction -- so we offer it as a diagnostic, not a cause. All evidence
is single-domain (anime) on one main model; whether the account transfers
to photographic identity or other DiT families is open, and contrary
locality findings in prior work may reflect that setting difference.

\section{Conclusion}\label{sec:conclusion}

We activated a frozen, pure-T2I anime diffusion transformer as a
one-shot reference-conditioned identity generator two independent, cheap
ways -- zero-training attention KV injection and a 0.06\%-parameter
band-LoRA -- each statistically tied with a same-class trained baseline
on identity (\texttt{strip} with IP-Adapter at zero training cost; the
band-LoRA with krea2 at 0.06\% of its parameters), reported as ties, not
wins, with a shuffled-reference control confirming the conditioning is
genuinely reference-driven. Using these mechanisms as probes, our central
finding is an anatomy: the one-shot identity signal in this frozen DiT is
\emph{distributed and composition-entangled}. Within the handles we could
test it is not localized -- to any layer band (a null replicated across
two held-out subject sets and concurring across both mechanisms), head
subset, or denoising phase -- and the collage defect is entangled with
identity rather than separable from it, as five matched-control repairs
each confirm by trading identity for coherence rather than removing the
defect. A peaked-versus-diffuse attention-selectivity signature, offered
as a cross-architecture diagnostic, further separates a model that
transfers structured identity from one that transfers only mean color.
Two measurement reversals along the way show that identity-match-only,
single-set evaluation is systematically optimistic for this class of
method. The composition-collapse mode is characterized but unfixed, and
geometric-attribute fidelity remains a residual failure even for the
recommended recipe.
